---
title: "Example of Fellegi-Sunter Match Probabilities"
author: "Lars Vilhuber"
date: "November 1, 2017"
output:
  html_document:
    highlight: monochrome
    keep_md: yes
    theme: journal
  pdf_document:
    highlight: monochrome
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
From the [INFO7470 lecture](https://www.vrdc.cornell.edu/info747x), we have
$$
R \equiv \frac{Pr [ \gamma_r | ab_r \in M]}{Pr[ \gamma_r | ab_r \in U]}
$$

In the lecture, John asks you to consider the case when the single variable under consideration is "sex":

$$
R \equiv \frac{Pr [ \gamma^1_r | ab_r \in M]}{Pr[ \gamma^1_r | ab_r \in U]}
$$
```{r setup_prob, echo=FALSE}
a_coding_error <- 0.01
b_coding_error <- 0
# from the source above:
p.year <- 2010
p_f <- 50.9
p_m <- 100-p_f
```

So assume that the binary variable $a_1 = b_1 =$ `sex` in this dataset is coded `m` for male and `f` for female, and that due to entry errors it may be miscoded in *A* about **`r a_coding_error * 100`%** of the time, but is recorded with **`r 100 - ( b_coding_error * 100)`%** correctly in *B*. Both datasets are drawn from and representative of the general U.S. population in 2010 (see f.i. [Age and Sex Composition: 2010](https://www.census.gov/prod/cen2010/briefs/c2010br-03.pdf)). 

# Question 1
What is $Pr [ \gamma^1_r | ab_r \in M]$?
```{r q1,echo=FALSE}
# we gloss over the configured b_coding_error for now
pr_g_m <- 1-a_coding_error
```

## Answer
When the two records are from the set of true matches, there is still a chance that the sex variable is miscoded on one of the two source records. Thus, up to **`r a_coding_error*100`%**, the error rate in *A*, regardless of sex, $Pr [ \gamma^1_r | ab_r \in M] =$ `r pr_g_m`.

# Question 2
What is $Pr [ \gamma^1_r | ab_r \in U]$?

## Answer:
In `r p.year`, `r p_f`% of the U.S. population  was female ($p_f$). So the likelihood that records in the set of non-matches have the same sex is approximately equal to the population percentage, if the samples are large enough. If *sex* =  `m`, then that probability is $p_m =$ **`r p_m`**, and if *sex* =  `f`, that probability is $p_f =$ **`r p_f`**. Thus
\begin{eqnarray}
Pr [ \gamma^1_r | ab_r \in U] & = & Pr [ \gamma^1_r | ab_r \in U , a = ``m"] Pr[ a = ``m"] \nonumber\\
                              &   & + Pr [ \gamma^1_r | ab_r \in U , a = ``f"] Pr[ a = ``f"] \nonumber\\
                              & = & p_m^2 + p_f^2 \nonumber\\
\end{eqnarray}
```{r, echo=FALSE}
pr_g_u <- (p_m/100)^2 + (p_f/100)^2
```
= `r pr_g_u`

# Question 3

What then is $w_r$?

##Answer

$w_r = log_2(R^*) = log_2($ `r (pr_g_m)` $/$ `r (pr_g_u)` $)  \approx log_2($ `r round((pr_g_m)/(pr_g_u),3)` $) \approx$  `r round(log2((pr_g_m)/(pr_g_u)),3)`.
